---
title: "analysis_tests"
output: html_document
---

```{r}
library(dplyr)
library(car)
library("lsmeans")
library(lme4)
library(viridis)
library(lmerTest)
library("ggplot2")
library(gridExtra)

BuRd <- c("#2066AC", "#4392C2", "#92C4DE", "#D1E5F0", "#F7F7F7", "#FDDBC6", "#F4A581", "#D5604C", "#B2182B")
BlueRed <- colorRampPalette(BuRd, interpolate="linear")
```

```{r}
#number of points to sample
npts <- c(36, 81, 144, 225, 324)
#sampling strategies
sampstrats <- c("rand", "grid", "trans", "envgeo")
#landscape dimensions (square)
dim = 40

#Create dataframe with all variable combos
params_full <- expand.grid(K = c(2, 4), 
                      phi = c(0.1, 0.5),
                      m = c(0.25, 1.0),
                      seed = c(1, 2, 3),
                      H = c(0.05 , 0.5),
                      r = c(0.3, 0.6),
                      it = 1:2,
                      npts = npts)

df <- params_full
df$stat <- df$K/10 + df$K*10 - df$m*10 + df$H*10 + df$r*10 + sample(df$it)/100 + sample(df$seed)/100 + npts

```


#Analysis

```{r}
#number of points to sample
npts <- c(36, 81, 144, 225, 324)
#sampling strategies
sampstrats <- c("rand", "grid", "trans", "envgeo")

#Create dataframe with all variable combos
params_full <- expand.grid(K = c(2, 4), 
                      phi = c(0.1, 0.5),
                      m = c(0.25, 1.0),
                      seed = c(1, 2, 3),
                      H = c(0.05 , 0.5),
                      r = c(0.3, 0.6),
                      npts = npts,
                      sampstrats = sampstrats,
                      it = 1:10)



df <- read.csv("LFMM_results.csv")
df[,colnames(params_full)] <- data.frame(lapply(df[,colnames(params_full)], as.factor))
df$stat <- df$TPR
```


```{r toy data}

params_toy <- expand.grid(K = c(0, 1), 
                      phi = c(0, 1),
                      m = c(1, 0),
                      seed = c(1, 1, 1),
                      H = c(0, 1),
                      r = c(0, 1),
                      npts = c(0.2,0.4,0.6,0.8,1),
                      sampstrats = c(0,0.75,0.25,1),
                      it=1:10)


range01 <- function(x){(x-min(x))/(max(x)-min(x))}

stat <-  0.4*params_toy$K + 0.1*params_toy$phi + 0.2*params_toy$m + 0.3*params_toy$H + 0.1*params_toy$r + 0.5*params_toy$npts + 0.1*params_toy$sampstrats + 0.5*sample(params_toy$it)/10
stat <- range01(stat)
?scale
stat <- jitter(stat, factor = 2)

df <- data.frame(lapply(params_full, as.factor))
df$stat <- stat
df$factornpts <- as.factor(params_full$npts)

df$npts <- params_full$npts
```


#summary plot
Continous summary plots (help visualize a plateau as npts increases) - this isn't done to compare the other variables because those are all categorical/it is easier to use the heatmaps rather than create a bajillion violin plots

```{r}

pK <- ggplot(df, aes(fill=K, y=stat, x=factor(npts))) + 
    geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis(discrete=T, option="plasma") +
    xlab("npts") +
    theme_bw()+
    geom_boxplot(width = 0.4, position = position_dodge(width = 0.9))+
    #geom_point(aes(col=K),position = position_dodge(width = 0.9), alpha=0.1)+
    scale_colour_viridis(discrete=T, option="plasma") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"))

pphi <- ggplot(df, aes(fill=phi, y=stat, x=factor(npts))) + 
    geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis(discrete=T, option="plasma") +
    xlab("npts") +
    theme_bw()+
    geom_boxplot(width = 0.4, position = position_dodge(width = 0.9))+
    #geom_point(aes(col=phi),position = position_dodge(width = 0.9), alpha=0.1)+
    scale_colour_viridis(discrete=T, option="plasma") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"))

pH <- ggplot(df, aes(fill=H, y=stat, x=factor(npts))) + 
    geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis(discrete=T, option="plasma") +
    xlab("npts") +
    theme_bw()+
    geom_boxplot(width = 0.4, position = position_dodge(width = 0.9))+
    #geom_point(aes(col=H),position = position_dodge(width = 0.9), alpha=0.1)+
    scale_colour_viridis(discrete=T, option="plasma") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"))

pr <- ggplot(df, aes(fill=r, y=stat, x=factor(npts))) + 
    geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis(discrete=T, option="plasma") +
    xlab("npts") +
    theme_bw()+
    geom_boxplot(width = 0.4, position = position_dodge(width = 0.9))+
    #geom_point(aes(col=r),position = position_dodge(width = 0.9), alpha=0.1)+
    scale_colour_viridis(discrete=T, option="plasma") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"))

pm <- ggplot(df, aes(fill=m, y=stat, x=factor(npts))) + 
    geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis(discrete=T, option="plasma") +
    xlab("npts") +
    theme_bw()+
    geom_boxplot(width = 0.4, position = position_dodge(width = 0.9))+
    scale_colour_viridis(discrete=T, option="plasma") +
    #geom_point(aes(col=m),position = position_dodge(width = 0.9), alpha=0.1)+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"))

ps <- ggplot(df, aes(fill=sampstrats, y=stat, x=factor(npts))) + 
    geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis(discrete=T, option="plasma") +
    xlab("npts") +
    theme_bw()+
    geom_boxplot(width = 0.4, position = position_dodge(width = 0.9))+
    #geom_point(aes(fill=sampstrats, col=sampstrats),position = position_dodge(width = 0.9), alpha=0.1)+
    scale_colour_viridis(discrete=T, option="plasma") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"))
  
```
```{r fig.height=7, fig.width=17}
grid.arrange(pK, pphi, pm, pH, pr, ps, nrow = 2)
```

# full model
```{r}
#Need to figure out if you need to subset this anova/remove n.s. variables at this stage or later
#mixed effect model
fullmod <- lmer(stat ~ npts + sampstrats + K + m + phi + H + r + (1 | seed), df)

#anova
aovmod <- anova(fullmod)

#fixed effects (minus intercept)
efmod <- fixef(fullmod)[-1]

#print result
aovmod
efmod
```

# model subset by sampling strategy
Determine how sampling strategy affects the relationships between the sim vars/npts

```{r fig.width=6, fig.height=4}
resdf <- data.frame()

for(s in sampstrats){
  subdf <- df[df$sampstrats == s, ]
  
  #Need to figure out if you need to subset this anova/remove n.s. variables at this stage or later
  #mixed effect model
  fullmod <- lmer(stat ~ npts + K + m + phi + H + r + (1 | seed), subdf)

  #anova to get p-values
  aovmod <- anova(fullmod)

  #fixed effects (minus intercept)
  efmod <- fixef(fullmod)[-1]
  
  #save results
  tempdf <- data.frame(fixef = efmod, p = aovmod$`Pr(>F)`)
  tempdf$param <- row.names(aovmod)
  tempdf$sampstrats <- s
  resdf <- rbind.data.frame(resdf, tempdf)
}


row.names(resdf) <- NULL
resdf$padj <- p.adjust(resdf$p, "fdr")

tempdf <- resdf
tempdf[tempdf$padj > 0.05, "fixef"] <- NA

p <- ggplot(tempdf, aes(param, sampstrats)) +
    geom_tile(aes(fill = fixef)) + 
    geom_text(aes(label = signif(fixef,  digits = 2), hjust = 0.5)) +
    scale_fill_gradient2(low = "#2066AC", mid="#F7F7F7", high = "#B2182B", midpoint = 0, limits=c(min(resdf$fixef),max(resdf$fixef))) +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          #legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 12),
          axis.text.y = element_text(color = "gray50", size = 12), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()

p

```


# model subset by npts and sampling strategy
Determine how sampling strategy/npts affects the affects of the sim vars


```{r real code, fig.width=12, fig.height=6}

resdf <- data.frame()

for(n in npts){
  for(s in sampstrats){
    subdf <- df[df$sampstrats == s & df$npts == n, ]
    
    #Need to figure out if you need to subset this anova/remove n.s. variables at this stage or later
    #mixed effect model
    fullmod <- lmer(stat ~ K + m + phi + H + r + (1 | seed), subdf)

    #anova to get p-values
    aovmod <- anova(fullmod)

    #fixed effects (minus intercept)
    efmod <- fixef(fullmod)[-1]
    
    #save results
    tempdf <- data.frame(fixef = fixef(fullmod)[-1], p = aovmod$`Pr(>F)`)
    tempdf$param <- row.names(aovmod)
    tempdf$npts <- n
    tempdf$sampstrats <- s
    resdf <- rbind.data.frame(resdf, tempdf)
  }
}

row.names(resdf) <- NULL
resdf$npts <- as.factor(resdf$npts)
resdf$padj <- p.adjust(resdf$p, "fdr")

## plot data
plts <- list()
for(i in 1:length(unique(resdf$param))){
  tempdf <- resdf[resdf$param == unique(resdf$param)[i],]
  tempdf[tempdf$padj > 0.05, "fixef"] <- NA
  p <- ggplot(tempdf, aes(npts, sampstrats)) +
    ggtitle(unique(tempdf$param)) +
    geom_tile(aes(fill = fixef)) + 
    geom_text(aes(label = round(fixef, digits = 2), hjust = 0.5)) +
    scale_fill_gradient2(low = "#2066AC", mid="#F7F7F7", high = "#B2182B", midpoint = 0, limits=c(min(resdf$fixef),max(resdf$fixef))) +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 12),
          axis.text.y = element_text(color = "gray50", size = 12), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()
  
  plts[[i]] <- p
}


do.call(grid.arrange, c(plts, nrow=2))

```

#summary stat plots by sampling strategy
```{r real code, fig.width=20, fig.height=6}

resdf <- data.frame()

for(n in npts){
  for(s in sampstrats){
    subdf <- df[df$sampstrats == s & df$npts == n, ]
    
    meandf <- data.frame()
    for(p in c("K", "m", "phi", "H", "r")){
      aggdf <- aggregate(subdf$stat, list(subdf[,p]), mean)
      aggdf[,1] <- paste(p,"=", aggdf[,1])
      aggdf$param <- p
      meandf <- rbind.data.frame(meandf, aggdf)
    }
    
    colnames(meandf) <- c("group", "mean", "param")
    
    #save results
    tempdf <- data.frame(meandf)
    tempdf$npts <- n
    tempdf$sampstrats <- s
    resdf <- rbind.data.frame(resdf, tempdf)
  }
}

row.names(resdf) <- NULL
resdf$npts <- as.factor(resdf$npts)

## plot data
plts <- list()
for(i in 1:length(unique(resdf$group))){
  tempdf <- resdf[resdf$group == unique(resdf$group)[i],]
  p <- ggplot(tempdf, aes(npts, sampstrats)) +
    ggtitle(unique(tempdf$group)) +
    geom_tile(aes(fill = mean)) + 
    geom_text(aes(label = round(mean, digits = 2), hjust = 0.5)) +
    scale_fill_viridis(limits=c(min(resdf$mean),max(resdf$mean)), option = "viridis") +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 14),
          axis.text.y = element_text(color = "gray50", size = 14), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()
  
  plts[[i]] <- p
}


plts2 <- list()
o <- 1
for(i in seq(1, length(plts), 2)){
  plts2[[o]] <- do.call(grid.arrange, c(plts[c(i,i+1)], nrow=2))
  o <- o+1
}

do.call(grid.arrange, c(plts2, nrow=1))


```

```{r}
#rainbow plot - not for use, but easier for my brain to interpret
## plot data
plts <- list()
for(i in 1:length(unique(resdf$group))){
  tempdf <- resdf[resdf$group == unique(resdf$group)[i],]
  p <- ggplot(tempdf, aes(npts, sampstrats)) +
    ggtitle(unique(tempdf$group)) +
    geom_tile(aes(fill = mean)) + 
    geom_text(aes(label = round(mean, digits = 2), hjust = 0.5)) +
    scale_fill_viridis(limits=c(min(resdf$mean),max(resdf$mean)), option = "turbo") +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 14),
          axis.text.y = element_text(color = "gray50", size = 14), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()
  
  plts[[i]] <- p
}


plts2 <- list()
o <- 1
for(i in seq(1, length(plts), 2)){
  plts2[[o]] <- do.call(grid.arrange, c(plts[c(i,i+1)], nrow=2))
  o <- o+1
}

do.call(grid.arrange, c(plts2, nrow=1))
```

