---
title: "GEA Analysis"
output: github_document
---

```{r, warning = FALSE, message = FALSE, include = FALSE}
library("here")
library("sjPlot")
library("dplyr")
library("lme4")
library("viridis")
library("lmerTest")
library("ggplot2")
library("gridExtra")
library("grid")
library("gt")
library("purrr")

source("analysis_functions.R")

p4path <- here(dirname(getwd()), "p4_analysis", "outputs")
p3path <- here(dirname(getwd()), "p3_methods", "outputs")
```

# 1. LFMM

## 1.1 Individual sampling

### 1.1.1 Summary plots

```{r, fig.width=16, fig.height=7, include = FALSE}
lfmm_ind <- format_lfmm(here(p3path, "lfmm_indsampling_results.csv")) 

lfmm_ind_tidy <- lfmm_ind %>% 
  filter(padj == "fdr") %>% 
  filter(sig == 0.05) %>% 
  group_by(method, K_selection) %>%
  group_split()

# GO BACK AND FIX K 
K_plots <- map(lfmm_ind_tidy, lfmm_plotter, stat = "K.1", colpal = "turbo")
TPR_plots <- map(lfmm_ind_tidy, lfmm_plotter, stat = "TPRCOMBO", colpal = "plasma", maxv = 0.17)
FDR_plots <- map(lfmm_ind_tidy, lfmm_plotter, stat = "FDRCOMBO", colpal = "viridis", direction = -1, maxv = 0.96)
TOTALN_plots <- map(lfmm_ind_tidy, lfmm_plotter, stat = "TOTALN", colpal = "viridis")

```

### K
```{r, fig.width=16, fig.height=7}
walk(K_plots, grid.arrange)
```

### TPRCOMBO
```{r, fig.width=16, fig.height=7}
walk(TPR_plots, grid.arrange)
```

### FDRCOMBO
```{r, fig.width=16, fig.height=7}
walk(FDR_plots, grid.arrange)
```
### TOTALN
```{r, fig.width=16, fig.height=7}
walk(TOTALN_plots, grid.arrange)
```

### 1.1.2 Model summaries
```{r}
lfmm_ind <- 
  lfmm_ind %>% 
  filter(padj == "fdr") %>%
  filter(alpha == 0.05) %>%
  filter(method == "ridge") %>%
  filter(K_selection == "tracy.widom")
```

```{r}
run_lmer(lfmm_ind, "TPRCOMBO", filepath = here(p4path, "LFMM_individual_TPR.csv"))
run_lmer(lfmm_ind, "FDRCOMBO", filepath = here(p4path, "LFMM_individual_FDR.csv"))
run_lmer(lfmm_ind, "TOTALN", filepath = here(p4path, "LFMM_individual_TOTALN.csv"))
```

### 1.1.3 Megaplots

```{r,fig.width = 21, fig.height = 13}
MEGAPLOT(lfmm_ind, "K.1", colpal = "turbo")
MEGAPLOT(lfmm_ind, "TPRCOMBO", colpal = "plasma")
MEGAPLOT(lfmm_ind, "FDRCOMBO", colpal = "viridis", direction = -1)
MEGAPLOT(lfmm_ind, "TOTALN", colpal = "viridis")
```
### 1.1.4 Latent factor versus sample number test

```{r, fig.width=16, fig.height=7, include = FALSE}
lfmm_ind <- format_lfmm(here(p3path, "lfmm_indsampling_fullK_results.csv")) 

lfmm_ind_tidy <- lfmm_ind %>% 
  filter(padj == "fdr") %>% 
  filter(sig == 0.05) %>% 
  group_by(method, K_selection) %>%
  group_split()

K_plots <- map(lfmm_ind_tidy, lfmm_plotter, stat = "K.1", colpal = "turbo")
TPR_plots <- map(lfmm_ind_tidy, lfmm_plotter, stat = "TPRCOMBO", colpal = "plasma")
FDR_plots <- map(lfmm_ind_tidy, lfmm_plotter, stat = "FDRCOMBO", colpal = "viridis", direction = -1)
TOTALN_plots <- map(lfmm_ind_tidy, lfmm_plotter, stat = "TOTALN", colpal = "viridis")
```

```{r, fig.width=16, fig.height=7}
walk(K_plots, grid.arrange)
walk(TPR_plots, grid.arrange)
walk(FDR_plots, grid.arrange)
walk(TOTALN_plots, grid.arrange)
```

## 1.2 Site sampling

### 1.2.1 Summary plots

```{r, fig.width=16, fig.height=7, include = FALSE}
lfmm_site <- format_lfmm(here(dirname(getwd()), "p3_methods", "outputs", "lfmm_sitesampling_results.csv")) 

lfmm_site_tidy <- lfmm_site %>% 
  filter(padj == "fdr") %>% 
  filter(alpha == 0.05) %>% 
  group_by(method, K_selection) %>%
  group_split()
```

```{r, fig.width=16, fig.height=7, include = FALSE}
# GO BACK AND FIX K 
K_plots <- map(lfmm_site_tidy, lfmm_plotter, stat = "K.1", colpal = "turbo")
TPR_plots <- map(lfmm_site_tidy, lfmm_plotter, stat = "TPRCOMBO", colpal = "plasma")
FDR_plots <- map(lfmm_site_tidy, lfmm_plotter, stat = "FDRCOMBO", colpal = "viridis", direction = -1)
TOTALN_plots <- map(lfmm_site_tidy, lfmm_plotter, stat = "TOTALN", colpal = "viridis")
```

### K
```{r, fig.width=16, fig.height=7}
walk(K_plots, grid.arrange)
```

### TPRCOMBO
```{r, fig.width=16, fig.height=7}
walk(TPR_plots, grid.arrange)
```

### FDRCOMBO
```{r, fig.width=16, fig.height=7}
walk(FDR_plots, grid.arrange)
```

### TOTALN
```{r, fig.width=16, fig.height=7}
walk(TOTALN_plots, grid.arrange)
```

### 1.1.2 Model summaries
```{r}
lfmm_site <- 
  lfmm_site %>% 
  filter(padj == "fdr") %>%
  filter(alpha == 0.05) %>%
  filter(method == "ridge") %>%
  filter(K_selection == "tracy.widom")
```

```{r}
run_lmer(lfmm_site, "TPRCOMBO", filepath = here(p4path, "LFMM_site_TPR.csv"))
run_lmer(lfmm_site, "FDRCOMBO", filepath = here(p4path, "LFMM_site_FDR.csv"))
run_lmer(lfmm_site, "TOTALN", filepath = here(p4path, "LFMM_site_TOTALN.csv"))
```

### 1.2.3 Megaplots

```{r,fig.width = 21, fig.height = 13}
MEGAPLOT(lfmm_site, "K.1", colpal = "turbo")
MEGAPLOT(lfmm_site, "TPRCOMBO", colpal = "plasma")
MEGAPLOT(lfmm_site, "FDRCOMBO", colpal = "viridis", direction = -1)
MEGAPLOT(lfmm_site, "TOTALN", colpal = "viridis")
```
### 1.2.4 Latent factor versus sample number test

```{r, fig.width=16, fig.height=7, include = FALSE}
lfmm_site <- format_lfmm(here(p3path, "lfmm_sitesampling_fullK_results.csv")) 

lfmm_site_tidy <- lfmm_site %>% 
  filter(padj == "fdr") %>% 
  filter(sig == 0.05) %>% 
  group_by(method, K_selection) %>%
  group_split()

K_plots <- map(lfmm_site_tidy, lfmm_plotter, stat = "K.1", colpal = "turbo")
TPR_plots <- map(lfmm_site_tidy, lfmm_plotter, stat = "TPRCOMBO", colpal = "plasma")
FDR_plots <- map(lfmm_site_tidy, lfmm_plotter, stat = "FDRCOMBO", colpal = "viridis", direction = -1)
TOTALN_plots <- map(lfmm_site_tidy, lfmm_plotter, stat = "TOTALN", colpal = "viridis", direction = -1)
```

```{r, fig.width=16, fig.height=7}

walk(K_plots, grid.arrange)
walk(TPR_plots, grid.arrange)
walk(FDR_plots, grid.arrange)
walk(TOTALN_plots, grid.arrange) 
```

# 2. RDA

## 2.1 Individual sampling

### 2.1.1 Summary plots

```{r, fig.width=16, fig.height=7}
rda_ind <- format_rda(here(dirname(getwd()), "p3_methods", "outputs", "rda_indsampling_results.csv"))
rda_ind_tidy <- 
  rda_ind %>%
  filter(padj == "fdr") %>%
  filter(sig == 0.05) %>%
  group_by(correctPC) %>%
  group_split()
```

```{r, fig.width=16, fig.height=7, include = FALSE}
TPR_plots <- map(rda_ind_tidy, rda_plotter, stat = "TPRCOMBO", colpal = "plasma", maxv = 0.088)
FDR_plots <- map(rda_ind_tidy, rda_plotter, stat = "FDRCOMBO", colpal = "viridis", direction = -1, maxv = 0.14)
TOTALN_plots <- map(rda_ind_tidy, rda_plotter, stat = "TOTALN", colpal = "viridis")
```

```{r, fig.width=16, fig.height=7}
walk(TPR_plots, grid.arrange)
```

```{r, fig.width=16, fig.height=7}
walk(FDR_plots, grid.arrange)
```

### 2.1.2 Model summaries
```{r}
rda_ind <- 
  rda_ind %>% 
  filter(padj == "fdr") %>%
  filter(sig == 0.05) %>%
  filter(correctPC == "FALSE") 
```

```{r}
run_lmer(rda_ind, "TPRCOMBO", filepath = here(p4path, "RDA_individual_TPR.csv"))
run_lmer(rda_ind, "FDRCOMBO", filepath = here(p4path, "RDA_individual_FDR.csv"))
run_lmer(rda_ind, "TOTALN", filepath = here(p4path, "RDA_individual_TOTALN.csv"))
```

### 1.1.3 Megaplots

```{r,fig.width = 21, fig.height = 13}
MEGAPLOT(rda_ind, "TPRCOMBO", colpal = "plasma")
MEGAPLOT(rda_ind, "FDRCOMBO", colpal = "viridis", direction = -1)
MEGAPLOT(rda_ind, "TOTALN", colpal = "viridis")
```

## 2.2 Site sampling

### 2.2.1 Summary plots

```{r, fig.width=16, fig.height=7}
rda_site <- format_rda(here(dirname(getwd()), "p3_methods", "outputs", "rda_sitesampling_results.csv")) 
rda_site_tidy <- 
  rda_site %>%
  filter(padj == "fdr") %>%
  filter(sig == 0.05) %>%
  group_by(correctPC) %>%
  group_split()
```

```{r, fig.width=16, fig.height=7, include = FALSE}
TPR_plots <- map(rda_site_tidy, rda_plotter, stat = "TPRCOMBO", colpal = "plasma", maxv = 0.066)
FDR_plots <- map(rda_site_tidy, rda_plotter, stat = "FDRCOMBO", colpal = "viridis", direction = -1, maxv = 0.092)
TOTALN_plots <- map(rda_site_tidy, rda_plotter, stat = "TOTALN", colpal = "viridis")
```

```{r, fig.width=16, fig.height=7}
walk(TPR_plots, grid.arrange)
```

```{r, fig.width=16, fig.height=7}
walk(FDR_plots, grid.arrange)
```

```{r, fig.width=16, fig.height=7}
walk(TOTALN_plots, grid.arrange)
```

### 2.2.2 Model summaries
```{r}
rda_site <- 
  rda_site %>% 
  filter(padj == "fdr") %>%
  filter(sig == 0.05) %>%
  filter(correctPC == "FALSE") 
```

```{r}
run_lmer(rda_site, "TPRCOMBO", filepath = here(p4path, "RDA_site_TPR.csv"))
run_lmer(rda_site, "FDRCOMBO", filepath = here(p4path, "RDA_site_FDR.csv"))
run_lmer(rda_site, "TOTALN", filepath = here(p4path, "RDA_site_TOTALN.csv"))
```

### 2.2.3 Megaplots

```{r,fig.width = 21, fig.height = 13}
MEGAPLOT(rda_site, "TPRCOMBO", colpal = "plasma")
MEGAPLOT(rda_site, "FDRCOMBO", colpal = "viridis", direction = -1)
MEGAPLOT(rda_site, "TOTALN", colpal = "viridis")
```
