---
title: "IBD/IBE Analysis"
output: github_document
---

```{r, warning=FALSE, message=FALSE, fig.width = 21, fig.height = 13}
library("here")
library("sjPlot")
library("dplyr")
library("lme4")
library("viridis")
library("lmerTest")
library("ggplot2")
library("gridExtra")
library("gt")
library("ggthemes")

source("analysis_functions.R")

p4path <- here(dirname(getwd()), "p4_analysis", "outputs")
p3path <- here(dirname(getwd()), "p3_methods", "outputs")
```

# 1. MMRR

## 1.1 Individual sampling

### 1.1.1 Summary plots

```{r, fig.width=16, fig.height=7}
mmrr_ind <- format_mmrr(here(p3path, "mmrr_indsampling_results.csv"))

# overall error
summary_hplot(mmrr_ind, "ratio_ae", colpal = "viridis", direction = -1)
summary_hplot(mmrr_ind, "env_ae", colpal = "viridis", direction = -1)
summary_hplot(mmrr_ind, "geo_ae", colpal = "viridis", direction = -1)

# bias
summary_hplot(mmrr_ind, "ratio_err", divergent = TRUE)
summary_hplot(mmrr_ind, "comboenv_err", divergent = TRUE)
summary_hplot(mmrr_ind, "geo_err", divergent = TRUE)
```

### 1.1.2 Model summaries

```{r}
run_lmer(mmrr_ind, "ratio_ae", filepath = here(p4path, "MMRR_individual_ratioerr.csv"))
run_lmer(mmrr_ind, "geo_ae", filepath = here(p4path, "MMRR_individual_geoerr.csv"))
run_lmer(mmrr_ind, "env_ae", filepath = here(p4path, "MMRR_individual_enverr.csv"))
```

### 1.1.3 Megaplots

```{r, fig.width = 21, fig.height = 13}
MEGAPLOT(mmrr_ind, "ratio_ae", colpal = "viridis", direction = -1)
MEGAPLOT(mmrr_ind, "ratio_err", colpal = "viridis", divergent = TRUE)
MEGAPLOT(mmrr_ind, "comboenv_err", divergent = TRUE)
MEGAPLOT(mmrr_ind, "geo_err", divergent = TRUE)
```

## 1.2 Site sampling

### 1.2.1 Summary plots

```{r, fig.width=16, fig.height=7}
mmrr_site <- format_mmrr(here(dirname(getwd()), "p3_methods", "outputs", "mmrr_sitesampling_results.csv"))

# overall error
summary_hplot(mmrr_site, "ratio_ae", colpal = "viridis", direction = -1)
summary_hplot(mmrr_site, "env_ae", colpal = "viridis", direction = -1)
summary_hplot(mmrr_site, "geo_ae", colpal = "viridis", direction = -1)

# bias
summary_hplot(mmrr_site, "ratio_err", divergent = TRUE)
summary_hplot(mmrr_site, "comboenv_err", divergent = TRUE)
summary_hplot(mmrr_site, "geo_err", divergent = TRUE)
```

### 1.2.2 Model summaries

```{r}
run_lmer(mmrr_site, "ratio_ae", filepath = here(p4path, "MMRR_site_ratioerr.csv"))
run_lmer(mmrr_site, "geo_ae", filepath = here(p4path, "MMRR_site_geoerr.csv"))
run_lmer(mmrr_site, "env_ae", filepath = here(p4path, "MMRR_site_enverr.csv"))
```

### 1.2.3 Megaplots

```{r, fig.width = 21, fig.height = 13}
MEGAPLOT(mmrr_site, "ratio_ae", colpal = "viridis", direction = -1)
MEGAPLOT(mmrr_site, "ratio_err", divergent = TRUE)
MEGAPLOT(mmrr_site, "comboenv_err", divergent = TRUE)
MEGAPLOT(mmrr_site, "geo_err", divergent = TRUE)
```

# 2. GDM

## 2.1 Individual sampling

### 2.1.1 Summary plots

```{r, fig.width=16, fig.height=7}
gdm_ind <- format_gdm(here(dirname(getwd()), "p3_methods", "outputs", "gdm_indsampling_results.csv"))

# overall error
summary_hplot(gdm_ind, "ratio_ae", colpal = "viridis", direction = -1)
summary_hplot(gdm_ind, "env_ae", colpal = "viridis", direction = -1)
summary_hplot(gdm_ind, "geo_ae", colpal = "viridis", direction = -1)

# bias
summary_hplot(gdm_ind, "ratio_err", divergent = TRUE)
summary_hplot(gdm_ind, "comboenv_err", divergent = TRUE)
summary_hplot(gdm_ind, "geo_err", divergent = TRUE)
```

### 2.1.2 Model summaries

```{r}
run_lmer(gdm_ind, "ratio_ae", filepath = here(p4path, "GDM_individual_ratioerr.csv"))
run_lmer(gdm_ind, "geo_ae", filepath = here(p4path, "GDM_individual_geoerr.csv"))
run_lmer(gdm_ind, "env_ae", filepath = here(p4path, "GDM_individual_enverr.csv"))
```

### 2.1.3 Megaplots

```{r, fig.width = 21, fig.height = 13}
MEGAPLOT(gdm_ind, "ratio_ae", colpal = "viridis", direction = -1)
MEGAPLOT(gdm_ind, "ratio_err", divergent = TRUE)
MEGAPLOT(gdm_ind, "comboenv_err", divergent = TRUE)
MEGAPLOT(gdm_ind, "geo_err", divergent = TRUE)
```

### 2.1.4 Prop NA

Confirm that the distribution of NAs is as expected and the proportions are small

```{r,  fig.width = 21, fig.height = 13}
MEGAPLOT(gdm_ind, "geo_coeff", aggfunc = "prop_na", colpal = "mako")
```

## 2.2 Site sampling

### 2.2.1 Summary plots

```{r, fig.width=16, fig.height=7}
gdm_site <- format_gdm(here(dirname(getwd()), "p3_methods", "outputs", "gdm_sitesampling_results.csv"))

# overall error
summary_hplot(gdm_site, "ratio_ae", colpal = "viridis", direction = -1)
summary_hplot(gdm_site, "env_ae", colpal = "viridis", direction = -1)
summary_hplot(gdm_site, "geo_ae", colpal = "viridis", direction = -1)

# bias
summary_hplot(gdm_site, "ratio_err", divergent = TRUE)
summary_hplot(gdm_site, "comboenv_err", divergent = TRUE)
summary_hplot(gdm_site, "geo_err", divergent = TRUE)
```

### 2.2.2 Model summaries

```{r}
run_lmer(gdm_site, "ratio_ae", filepath = here(p4path, "GDM_site_ratioerr.csv"))
run_lmer(gdm_site, "geo_ae", filepath = here(p4path, "GDM_site_geoerr.csv"))
run_lmer(gdm_site, "env_ae", filepath = here(p4path, "GDM_site_enverr.csv"))
```

### 2.2.3 Megaplots

```{r, fig.width = 21, fig.height = 13}
MEGAPLOT(gdm_site, "ratio_ae", colpal = "viridis", direction = -1)
MEGAPLOT(gdm_site, "ratio_ae", colpal = "viridis", direction = -1)
MEGAPLOT(gdm_site, "ratio_ae", colpal = "viridis", direction = -1)

MEGAPLOT(gdm_site, "ratio_err", divergent = TRUE)
MEGAPLOT(gdm_site, "comboenv_err", divergent = TRUE)
MEGAPLOT(gdm_site, "geo_err", divergent = TRUE)
```

### 2.2.4 Prop NA

Confirm that the distribution of NAs is as expected and the proportions are small

```{r,  fig.width = 21, fig.height = 13}
MEGAPLOT(gdm_site, "geo_coeff", aggfunc = "prop_na", colpal = "mako")
```

# 3. Comparison of error distirbutions for MMRR and GDM


```{r, fig.width = 5, fig.height = 5}
mmrr_ind %>%
  filter(sampstrat != "full") %>%
  ggplot(aes(x = ratio_ae, fill = m, colour = m)) +
  geom_density(alpha = 0.5) +
  theme_few() +
  scale_fill_viridis_d(direction = -1, end = 0.7, begin = 0.3, option = "mako") +
  scale_color_viridis_d(direction = -1, end = 0.7, begin = 0.3, option = "mako")


gdm_ind %>%
  filter(sampstrat != "full") %>%
  ggplot(aes(x = ratio_ae, fill = m, colour = m)) +
  geom_density(alpha = 0.5) +
  theme_few() +
  scale_fill_viridis_d(direction = -1, end = 0.7, begin = 0.3, option = "mako") +
  scale_color_viridis_d(direction = -1, end = 0.7, begin = 0.3, option = "mako")
```

# 4. Compare results of MMRR and GDM

```{r}
plot(mmrr_ind$geo_coeff, gdm_ind$geo_coeff, col = gdm_ind$m)
legend("topleft", c("0.25", "1"), col = c("black", "red"), pch = 1)
plot(mmrr_ind$comboenv_coeff, gdm_ind$comboenv_coeff)

df <- data.frame(mmrr_ind, 
                 geo_mg = gdm_ind$geo_coeff/mmrr_ind$geo_coeff, 
                 env_mg = gdm_ind$comboenv_coeff/mmrr_ind$comboenv_coeff,
                 ratio_mg = gdm_ind$ratio/mmrr_ind$ratio )
```

```{r, fig.width=16, fig.height=7}
summary_hplot(df, "geo_mg")
summary_hplot(df, "env_mg")
summary_hplot(df, "ratio_mg")
```

```{r, fig.width = 21, fig.height = 13}
MEGAPLOT(df, "geo_mg")
MEGAPLOT(df, "env_mg")
MEGAPLOT(df, "ratio_mg")
```

```{r}
plot(mmrr_site$geo_coeff, gdm_site$geo_coeff, col = gdm_site$m)
legend("topleft", c("0.25", "1"), col = c("black", "red"), pch = 1)
plot(mmrr_site$comboenv_coeff, gdm_site$comboenv_coeff)

df <- data.frame(mmrr_site, 
                 geo_mg = gdm_site$geo_coeff/mmrr_site$geo_coeff, 
                 env_mg = gdm_site$comboenv_coeff/mmrr_site$comboenv_coeff,
                 ratio_mg = gdm_site$ratio/mmrr_site$ratio )
```

```{r, fig.width=16, fig.height=7}
summary_hplot(df, "geo_mg")
summary_hplot(df, "env_mg")
summary_hplot(df, "ratio_mg")
```
