---
title: "analysis_tests"
output: html_document
---


```{r}
library("dplyr")
library("car")
library("lsmeans")
library("lme4")
library("viridis")
library("lmerTest")
library("ggplot2")
library("gridExtra")
library("MuMIn")
library("emmeans")
library("effects")

source("analysis_functions.R")
BuRd <- c("#2066AC", "#4392C2", "#92C4DE", "#D1E5F0", "#F7F7F7", "#FDDBC6", "#F4A581", "#D5604C", "#B2182B")
BlueRed <- colorRampPalette(BuRd, interpolate="linear")
```


#Analysis

```{r}
#number of sites to sample
nsamp <- c(9, 16, 25, 2000)
#sampling strategies
sampstrat <- c("rand", "envgeo", "equi","full")

#Create dataframe with all variable combos
params_full <- expand.grid(K = c(2, 4), 
                      phi = c(0.1, 0.5),
                      m = c(0.25, 1.0),
                      seed = c(1, 2, 3),
                      H = c(0.05 , 0.5),
                      r = c(0.3, 0.6),
                      nsamp = nsamp,
                      sampstrat = sampstrat,
                      it = 0:9)

df <- read.csv("/Users/Anusha/Documents/GitHub/LandGenSamp/p3_methods/outputs/gdm_sitesampling_results.csv")

df[,colnames(params_full)] <- data.frame(lapply(df[,colnames(params_full)], as.factor))
summary(df)

colnames(df)[13:15] <- c( "env1_err",  "env2_err", "geo_err")
df$null <- df$geo_coeff == "NULL" & df$env1_coeff == "NULL" & df$env2_coeff == "NULL"
#df <- df[df$env1_coeff != "NULL",]
#df <- df[df$env2_coeff != "NULL",]
#df <- df[df$geo_coeff != "NULL",]

df$geo_err <- as.numeric(df$geo_err)
df$env1_err <- as.numeric(df$env1_err)
df$env2_err <- as.numeric(df$env2_err)

df$geo_coeff <- as.numeric(df$geo_coeff)
df$env1_coeff <- as.numeric(df$env1_coeff)
df$env2_coeff <- as.numeric(df$env2_coeff)

df$comboenv_coeff <- df$env1_coeff + df$env2_coeff
df$comboenv_err <- (df$env1_err + df$env2_err)/2


#REMOVE FULL DATA FOR MODELS
moddf <- df[df$sampstrat != "full",]

```

```{r}
#Create dataframe with all variable combos
params <- expand.grid(K = c(1, 2), 
                      phi = c(0.1, 0.5),
                      m = c(0.25, 1.0),
                      seed = c(1, 2, 3),
                      H = c(0.05, 0.5),
                      r = c(0.3, 0.6),
                      it = 0:9)

 cdf <- data.frame()
for(i in 1:nrow(params)){
  p <- params[i,]
  subdf <- df[df$K == p$K & df$phi == p$phi & df$m == p$m & df$seed == p$seed & df$H == p$H & df$r == p$r & df$it == p$it, ]
  full <- subdf[subdf$sampstrat == "full",]
  sub <- subdf[subdf$sampstrat != "full",]
  sub$geo_mpe <-sub$geo_err/full$geo_coeffn
  actual <- sub$env1_coeffn + sub$env2_coeffn
  forecast <- full$env1_coeffn + full$env2_coeffn
  sub$comboenv_mpe <- abs(actual - forecast)/actual
  cdf <- rbind(cdf, sub)
}

```


```{r fig.width=20, fig.height=7}
cdf <- data.frame()
for(i in 1:nrow(params)){
  p <- params[i,]
  subdf <- df[df$K == p$K & df$phi == p$phi & df$m == p$m & df$seed == p$seed & df$H == p$H & df$r == p$r & df$it == p$it, ]
  full <- subdf[subdf$sampstrat == "full",]
  sub <- subdf[subdf$sampstrat != "full",]
  sub$ratio <- sub$comboenv_coeff/sub$geo_coeff
  full$ratio <-full$comboenv_coeff/full$geo_coeff
  sub$ratio_rmse <- abs(sub$ratio - full$ratio)
  cdf <- rbind(cdf, sub)
}

summary_hplot(cdf, "geo_err")
summary_hplot(cdf, "comboenv_err")

summary_hplot(cdf, "ratio_rmse")

summary_hplot(cdf, "null")

```

#summary plots

```{r fig.width = 30, fig.height=15}
MEGAPLOT(moddf, "geo_err", option = "turbo", maxv = 3, na.rm = TRUE)
MEGAPLOT(moddf, "comboenv_err", option = "turbo", maxv = 3, na.rm = TRUE)
```




```{r fig.width = 30, fig.height=15}
MEGAPLOT(moddf, "ratio", option = "turbo", maxv = 0.3,  na.rm=TRUE)
```

```{r fig.width = 30, fig.height=15}

maxv = 0.85
MEGAPLOT(moddf, "comboenv_err", aggfunc = "prop_na", option = "turbo", maxv = maxv,  na.rm=TRUE)
MEGAPLOT(moddf, "geo_err", aggfunc = "prop_na", option = "turbo", maxv = maxv,  na.rm=TRUE)
MEGAPLOT(moddf, "geo_coeff", aggfunc = "prop_na", option = "turbo", maxv = maxv,  na.rm=TRUE)

MEGAPLOT(moddf, "env1_coeff", aggfunc = "prop_na", option = "turbo", maxv = maxv,  na.rm=TRUE)
MEGAPLOT(moddf, "env2_coeff", aggfunc = "prop_na", option = "turbo", maxv = maxv,  na.rm=TRUE)
```


```{r fig.width = 30, fig.height=15}
MEGAPLOT(moddf, "geo_coeffn", option = "turbo", max = 4, na.rm=TRUE)
MEGAPLOT(moddf, "comboenv_coeffn", option = "turbo", max = 0.5, na.rm=TRUE)

MEGAPLOT(moddf, "env1_coeffn", option = "turbo", max = 0.3, na.rm=TRUE)
MEGAPLOT(moddf, "env2_coeffn", option = "turbo", max = 0.3, na.rm=TRUE)
```

```{r fig.width=20, fig.height=7}
summary_vplot(moddf, "geo_coeff")
summary_hplot(df, "geo_err")

summary_hplot(df, stat_name = "geo_err", aggfunc = "mean", colpal = "viridis", direction = -1, na.rm=TRUE)
```

```{r fig.width=20, fig.height=7}
summary_vplot(df, "comboenv_err")
summary_hplot(df, "comboenv_err")

summary_hplot(df, "comboenv_err", aggfunc = "sd")
summary_hplot(df, "comboenv_err", aggfunc = "mean", colpal = "viridis", direction = -1)
```
```{r fig.width=20, fig.height=7}
summary_hplot(df, "ratio", aggfunc = "sd")
summary_hplot(df, "ratio", aggfunc = "mean", colpal = "viridis", direction = -1)
summary_hplot(df, "ratio", aggfunc = "prop_na", colpal = "viridis", direction = -1)
```
```{r fig.width=20, fig.height=7}
summary_hplot(df, "null", aggfunc = "mean", colpal = "viridis", direction = -1)
```

#VARIANCE EXPLORATION
```{r}
df$stat <- df$geo_err

df %>%
  group_by(seed) %>%
  summarise(var = var(geo_err, na.rm=TRUE)) %>%
  arrange(var)


ggplot(df, aes(fill=seed, y=geo_err, x=factor(seed))) + 
     #geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
     scale_fill_viridis(discrete=T, option="viridis") +
     theme_bw()+
     geom_point(aes(col=seed),position = position_jitterdodge(dodge.width = 0.9, jitter.width=0.9), alpha=0.1)+
     geom_boxplot(width = 0.4, position = position_dodge(width = 0.9), alpha = 0.5)+
     scale_colour_viridis(discrete=T, option="viridis") +
     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
           panel.background = element_blank(), axis.line = element_line(colour = "black"), text=element_text(size=21))

ggplot(df, aes(fill=seed, y=log(geo_err), x=factor(seed))) + 
     #geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
     scale_fill_viridis(discrete=T, option="viridis") +
     theme_bw()+
     geom_point(aes(col=seed),position = position_jitterdodge(dodge.width = 0.9, jitter.width=0.9), alpha=0.1)+
     geom_boxplot(width = 0.4, position = position_dodge(width = 0.9), alpha = 0.5)+
     scale_colour_viridis(discrete=T, option="viridis") +
     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
           panel.background = element_blank(), axis.line = element_line(colour = "black"), text=element_text(size=21))
```




# full model w/ interactions
```{r}
#Need to figure out if you need to subset this anova/remove n.s. variables at this stage or later
#mixed effect model
#na.action needs to be set to na.fail for dredge to work
fullmod <- lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + 
                    K*m + K*phi + K*H + K*r +
                    m*phi + m*H + m*r + 
                    phi*H + phi*r + 
                    H*r +
                    (1 | seed), moddf, na.action = "na.fail")

#anova
aovmod <- anova(fullmod)

#fixed effects (minus intercept)
efmod <- fixef(fullmod)[-1]

#print result
aovmod
efmod
```

```{r}
fullmod <-  lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + 
                    K*m + K*phi + K*H + 
                    m*phi + m*H + m*r + 
                    phi*H + phi*r + 
                    H*r +
                    (1 | seed), moddf, na.action = "na.fail")
combos <- dredge(fullmod)
combos
```


```{r}
finalmod <- lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + 
                    K*m + K*phi + K*H + 
                    m*phi + m*H + m*r + 
                    phi*H + phi*r + 
                    H*r +
                    (1 | seed), moddf, na.action = "na.fail")
#anova
aovmod <- anova(finalmod)
aovmod

```


```{r}
ints <- c("K*phi", "K*H", "K*m", 
          "m*phi", "m*H", "m*r",
          "phi*H", "phi*r", 
          "H*r")

effplts <- list()
for(i in 1:length(ints)){
    eff <- effect(ints[i], finalmod)
    effplts[[i]] <- eff
    print(plot(eff))
}


par(mfrow=c(2,4))
for(i in 1:length(effplts)){
   p <- plot(effplts[[i]])
   print(p)
}



effplts2 <- list()

for(i in 1:length(effplts)){
    eff <- as.data.frame(effplts[[i]])
    v1 <- eff[,1]
    v2 <- eff[,2]
    p <- ggplot_gtable(ggplot_build(ggplot(eff, aes(factor(v1), linetype=factor(v2),
                                     color = factor(v2))) +
        geom_line(aes(y = fit, group=factor(v2)), size=1) +
        geom_ribbon(aes(ymin = lower, ymax = upper, group = factor(v2), fill = factor(v2), col = factor(v2)), 
                  alpha=0.2, 
                  linetype="dashed") + 
        xlab(colnames(eff)[1]) +
        ylab("") +
        scale_colour_manual(values = c("#756fd6","#30D5C8")) +
        scale_fill_manual(values = c("#756fd6", "#30D5C8")) +
        scale_x_discrete(expand = c(0.05, 0.05)) +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
              panel.background = element_blank(), axis.line = element_line(colour = "black"), text=element_text(size=21),
              plot.margin=unit(rep(0.4,4),"cm")) +
        guides(col=guide_legend(title=colnames(eff)[2], labels = levels(factor(v2))), linetype = FALSE, fill = FALSE, size = FALSE)))
    
    effplts2[[i]] <- p
    gc()

}




```


```{r fig.width = 10, fig.height = 14}
do.call(grid.arrange, c(effplts2, ncol=2))
```


# full model w/o interactions
```{r}
#Need to figure out if you need to subset this anova/remove n.s. variables at this stage or later
#mixed effect model
#na.action needs to be set to na.fail for dredge to work
fullmod <- lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + 
                    (1 | seed), moddf, na.action = "na.fail")

#anova
aovmod <- anova(fullmod)

#fixed effects (minus intercept)
efmod <- fixef(fullmod)[-1]

#print result
aovmod
efmod
```
```{r}
combos <- dredge(fullmod)
combos
```

```{r}
combos <- dredge(fullmod, rank = "BIC")
combos
```



```{r pairwise comparison of npts, fig.width=5, fig.height=5}

model <- lmer(stat ~ nsamp + K + m + phi + H + r + (1 | seed), moddf, na.action = "na.fail")

pw <- emmeans(model, list(pairwise ~ nsamp), adjust = "tukey")

resdf <- data.frame(pw$`pairwise differences of nsamp`)
resdf$p1 <- gsub("\\ -.*", "", resdf$X1)
resdf$p2 <- gsub(".*- ", "", resdf$X1)

resdf$estimate[resdf$p.value > 0.05] <- NA

resdf$p1 <- factor(resdf$p1, ordered=TRUE, levels = as.character(nsamp))
resdf$p2 <- factor(resdf$p2, ordered=TRUE, levels = as.character(nsamp))

ggplot(resdf, aes(p1,p2)) +
    geom_tile(aes(fill = estimate)) + 
    geom_text(aes(label = signif(estimate,  digits = 2), hjust = 0.5)) +
    scale_fill_gradient2(low = "#2066AC", mid="#F7F7F7", high = "#B2182B", midpoint = 0, limits=c(min(resdf$estimate),max(resdf$estimate))) +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          #legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 12),
          axis.text.y = element_text(color = "gray50", size = 12), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()

```
```{r pairwise comparison of sampstrat, fig.width=10, fig.height=10}
#TukeyHSD

model <- lmer(log(stat) ~ nsamp + sampstrat + K + m + phi + H + r + (1 | seed), moddf, na.action = "na.fail")

pw <- emmeans(model, list(pairwise ~ sampstrat), adjust = "tukey")

resdf <- data.frame(pw$`pairwise differences of sampstrat`)
resdf$p1 <- gsub("\\ -.*", "", resdf$X1)
resdf$p2 <- gsub(".*- ", "", resdf$X1)


resdf$estimate[resdf$p.value > 0.05] <- NA

resdf$p1 <- factor(resdf$p1, ordered=TRUE, levels = as.character(c("envgeo","grid","rand","trans")))
resdf$p2 <- factor(resdf$p2, ordered=TRUE, levels = as.character(c("envgeo","grid","rand","trans")))

ggplot(resdf, aes(p1,p2)) +
    geom_tile(aes(fill = estimate)) + 
    geom_text(aes(label = signif(estimate,  digits = 2), hjust = 0.5)) +
    scale_fill_gradient2(low = "#2066AC", mid="#F7F7F7", high = "#B2182B", midpoint = 0, limits=c(min(resdf$estimate),max(resdf$estimate))) +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          #legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 12),
          axis.text.y = element_text(color = "gray50", size = 12), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()

ggplot(moddf, aes(fill=sampstrat, y=log(stat), x=factor(nsamp))) + 
      #geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
      scale_fill_viridis(discrete=T, option="mako") +
      xlab("nsamp") +
      theme_bw()+
    geom_point(aes(fill=sampstrat, col=sampstrat), position = position_jitterdodge(dodge.width = 0.9, jitter.width = 0.5), alpha=0.2, size = 1)+
      geom_boxplot(width = 0.4, position = position_dodge(width = 0.9), alpha = 0.8)+
      scale_colour_viridis(discrete=T, option="mako") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line = element_line(colour = "black"), text=element_text(size=21))

ggplot(moddf, aes(fill=sampstrat, y=stat, x=factor(nsamp))) + 
      #geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
      scale_fill_viridis(discrete=T, option="mako") +
      xlab("nsamp") +
      theme_bw()+
    geom_point(aes(fill=sampstrat, col=sampstrat), position = position_jitterdodge(dodge.width = 0.9, jitter.width = 0.5), alpha=0.2, size = 1)+
      geom_boxplot(width = 0.4, position = position_dodge(width = 0.9), alpha = 0.8)+
      scale_colour_viridis(discrete=T, option="mako") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line = element_line(colour = "black"), text=element_text(size=21))



```

#which is best when?
```{r, fig.height=8, fig.width=5}
nsamp_sampstrat_mods(df, sampstrat = sampstrat, nsamp = nsamp)
```

```{r fig.width=10, fig.height=10}
sampstrat_nsamp_mods(df, sampstrat = sampstrat, nsamp = nsamp)
```

# model subset by sampling strategy
Determine how sampling strategy affects the relationships between the sim vars/nsamp

```{r fig.width=5, fig.height=4}
sampstrat_mod(df, sampstrat = sampstrat, full = TRUE)
```
```{r fig.width=5, fig.height=4}
nsamp_mod(df, nsamp = nsamp, full = TRUE)
```
# model subset by nsamp and sampling strategy
Determine how sampling strategy/nsamp affects the affects of the sim vars

```{r real code, fig.width=20, fig.height=4}
sampstrat_nsamp_mod_params(df, sampstrat = sampstrat, nsamp = nsamp, alpha = 0.10)
```


