---
title: "analysis_tests"
output: html_document
---


```{r}
library("dplyr")
library("car")
library("lsmeans")
library("lme4")
library("viridis")
library("lmerTest")
library("ggplot2")
library("gridExtra")
library("MuMIn")
library("emmeans")
library("effects")

source("sitesampling_analysis_functions.R")
BuRd <- c("#2066AC", "#4392C2", "#92C4DE", "#D1E5F0", "#F7F7F7", "#FDDBC6", "#F4A581", "#D5604C", "#B2182B")
BlueRed <- colorRampPalette(BuRd, interpolate="linear")

```


#Analysis

```{r}
#number of sites to sample
nsamp <- c(9, 16, 25, 2000)
#sampling strategies
sampstrat <- c("rand", "grid", "envgeo", "equi","full")

#Create dataframe with all variable combos
params_full <- expand.grid(K = c(2, 4), 
                      phi = c(0.1, 0.5),
                      m = c(0.25, 1.0),
                      seed = c(1, 2, 3),
                      H = c(0.05 , 0.5),
                      r = c(0.3, 0.6),
                      nsamp = nsamp,
                      sampstrat = sampstrat,
                      it = 0:9)

#df <- read.csv("/Users/Anusha/Documents/GitHub/LandGenSamp/p3_methods/sitesampling/outputs/LFMM/LFMM_sitesampling_results.csv")

df <- do.call(rbind, lapply(list.files(path = "/Users/Anusha/Documents/GitHub/LandGenSamp/p3_methods/sitesampling/outputs/LFMM/", full.names = TRUE, pattern="*.csv"), read.csv))

df[,colnames(params_full)] <- data.frame(lapply(df[,colnames(params_full)], as.factor))


#CORRECT TPRCOMBO CALC
#SEE NOTION NOTES (accidentally calculated non-unique loci)
df$TPRCOMBO <- (df$TPR1COMBO + df$TPR2COMBO)/2

#IMPORTANT NOTE ABOUT LFMM DATA:
#CURRENTLY IF THERE WERE NOT LOCI IDENTIFIED AS SIGNIFICANT TPR=0 AND FDR = NA (BECAUSE DIVIDE BY 0 = INF = NA). 
#NOT SURE HOW TO DEAL WITH THIS (I.E. LEAVE AS NA AND EXCLUDE FROM CALCULATIONS OR CONVERT NAS TO 0)
#I THINK CONVERTING FROM NA TO 0 IS THE MOST ACCURATE (SINCE THE FDR IS 0)
df$FDRCOMBO[df$TPRCOMBO == 0 & is.na(df$FDRCOMBO)] <- 0
df$FDR1COMBO[df$TPR1COMBO == 0 & is.na(df$FDR1COMBO)] <- 0
df$FDR2COMBO[df$TPR2COMBO == 0 & is.na(df$FDR2COMBO)] <- 0
df$FDR1[df$TPR1 == 0 & is.na(df$FDR1)] <- 0
df$FDR2[df$TPR2 == 0 & is.na(df$FDR2)] <- 0


#REMOVE FULL DATA FOR MODELS
moddf <- df[df$sampstrat != "full",]



```

```{r}
range(df$FDRCOMBO)
range(df$TPRCOMBO)

```
```{r}
summary(df)
```

```{r fig.width = 30, fig.height=40}
MEGAPLOT(moddf, "TPRCOMBO", max = 0.9, option = "turbo")
MEGAPLOT(moddf, "FDRCOMBO")
```



#EXPLORATORY

```{r fig.width=20, fig.height=8}
df$stat <- df$TPRCOMBO
summary_vplot(df)
summary_hplot(df)


df$stat <- log(df$TPRCOMBO+1)
summary_vplot(df)
summary_hplot(df)

df$stat <- sqrt(df$TPRCOMBO)
summary_vplot(df)
summary_hplot(df)
```


#TPRCOMBO
```{r fig.width=20, fig.height=7}
df$stat <- df$TPRCOMBO
moddf <- df[df$sampstrat != "full",]
summary_vplot(df)
summary_hplot(df)
```

# full model w/ interactions
```{r}
#Need to figure out if you need to subset this anova/remove n.s. variables at this stage or later
#mixed effect model
#na.action needs to be set to na.fail for dredge to work
fullmod <- lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + 
                    K*m + K*phi + K*H + K*r +
                    m*phi + m*H + m*r + 
                    phi*H + phi*r + 
                    H*r +
                    (1 | seed), moddf, na.action = "na.fail")

#anova
aovmod <- anova(fullmod)

#fixed effects (minus intercept)
efmod <- fixef(fullmod)[-1]

#print result
aovmod
efmod
```

```{r}
fullmod <-  lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + 
                    K*m + K*phi + K*H + 
                    m*phi + m*H + m*r + 
                    phi*H + phi*r + 
                    H*r +
                    (1 | seed), moddf, na.action = "na.fail")
combos <- dredge(fullmod)
combos
```


```{r}
finalmod <- lmer(stat ~ nsamp + K + m + phi + H + r +
                    K*phi + K*H + 
                    m*phi + m*H + m*r + 
                    phi*H + phi*r + 
                    H*r +
                    (1 | seed), moddf, na.action = "na.fail")
#anova
aovmod <- anova(finalmod)
aovmod

```


```{r}
ints <- c("K*phi", "K*H", 
          "m*phi", "m*H", "m*r",
          "phi*H", "phi*r", 
          "H*r")

effplts <- list()
for(i in 1:length(ints)){
    eff <- effect(ints[i], finalmod)
    effplts[[i]] <- eff
    print(plot(eff))
}


par(mfrow=c(2,4))
for(i in 1:length(effplts)){
   p <- plot(effplts[[i]])
   print(p)
}



effplts2 <- list()

for(i in 1:length(effplts)){
    eff <- as.data.frame(effplts[[i]])
    v1 <- eff[,1]
    v2 <- eff[,2]
    p <- ggplot_gtable(ggplot_build(ggplot(eff, aes(factor(v1), linetype=factor(v2),
                                     color = factor(v2))) +
        geom_line(aes(y = fit, group=factor(v2)), size=1) +
        geom_ribbon(aes(ymin = lower, ymax = upper, group = factor(v2), fill = factor(v2), col = factor(v2)), 
                  alpha=0.2, 
                  linetype="dashed") + 
        xlab(colnames(eff)[1]) +
        ylab("") +
        scale_colour_manual(values = c("#756fd6","#30D5C8")) +
        scale_fill_manual(values = c("#756fd6", "#30D5C8")) +
        scale_x_discrete(expand = c(0.05, 0.05)) +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
              panel.background = element_blank(), axis.line = element_line(colour = "black"), text=element_text(size=21),
              plot.margin=unit(rep(0.4,4),"cm")) +
        guides(col=guide_legend(title=colnames(eff)[2], labels = levels(factor(v2))), linetype = FALSE, fill = FALSE, size = FALSE)))
    
    effplts2[[i]] <- p
    gc()

}




```


```{r fig.width = 10, fig.height = 14}
do.call(grid.arrange, c(effplts2, ncol=2))
```


# full model w/o interactions
```{r}
#Need to figure out if you need to subset this anova/remove n.s. variables at this stage or later
#mixed effect model
#na.action needs to be set to na.fail for dredge to work
fullmod <- lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + 
                    (1 | seed), moddf, na.action = "na.fail")

moddf[moddf$sampstrat == "test",]
levels(moddf$sampstrat)
#anova
aovmod <- anova(fullmod)

#fixed effects (minus intercept)
efmod <- fixef(fullmod)[-1]

#print result
aovmod
efmod
```
```{r}
combos <- dredge(fullmod)
combos
```

```{r}
combos <- dredge(fullmod, rank = "BIC")
combos
```



```{r pairwise comparison of npts, fig.width=5, fig.height=5}

model <- lmer(stat ~ nsamp + K + m + phi + H + r + (1 | seed), moddf, na.action = "na.fail")

pw <- emmeans(model, list(pairwise ~ nsamp), adjust = "tukey")

resdf <- data.frame(pw$`pairwise differences of nsamp`)
resdf$p1 <- gsub("\\ -.*", "", resdf$X1)
resdf$p2 <- gsub(".*- ", "", resdf$X1)

resdf$estimate[resdf$p.value > 0.05] <- NA

resdf$p1 <- factor(resdf$p1, ordered=TRUE, levels = as.character(nsamp))
resdf$p2 <- factor(resdf$p2, ordered=TRUE, levels = as.character(nsamp))

ggplot(resdf, aes(p1,p2)) +
    geom_tile(aes(fill = estimate)) + 
    geom_text(aes(label = signif(estimate,  digits = 2), hjust = 0.5)) +
    scale_fill_gradient2(low = "#2066AC", mid="#F7F7F7", high = "#B2182B", midpoint = 0, limits=c(min(resdf$estimate),max(resdf$estimate))) +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          #legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 12),
          axis.text.y = element_text(color = "gray50", size = 12), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()

```


```{r pairwise comparison of sampstrat, fig.width=5, fig.height=5}
#TukeyHSD

model <- lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + (1 | seed), moddf, na.action = "na.fail")

pw <- emmeans(model, list(pairwise ~ sampstrat), adjust = "tukey")

resdf <- data.frame(pw$`pairwise differences of sampstrat`)
resdf$p1 <- gsub("\\ -.*", "", resdf$X1)
resdf$p2 <- gsub(".*- ", "", resdf$X1)


resdf$estimate[resdf$p.value > 0.05] <- NA

resdf$p1 <- factor(resdf$p1, ordered=TRUE, levels = as.character(c("envgeo","grid","rand","equi")))
resdf$p2 <- factor(resdf$p2, ordered=TRUE, levels = as.character(c("envgeo","grid","rand","equi")))

ggplot(resdf, aes(p1,p2)) +
    geom_tile(aes(fill = estimate)) + 
    geom_text(aes(label = signif(estimate,  digits = 2), hjust = 0.5)) +
    scale_fill_gradient2(low = "#2066AC", mid="#F7F7F7", high = "#B2182B", midpoint = 0, limits=c(min(resdf$estimate),max(resdf$estimate))) +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          #legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 12),
          axis.text.y = element_text(color = "gray50", size = 12), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()

```

#which is best when?
```{r, fig.height=8, fig.width=5}
nsamp_sampstrat_mods(df, sampstrat = sampstrat, nsamp = nsamp)
```

```{r fig.width=10, fig.height=10}
sampstrat_nsamp_mods(df, sampstrat = sampstrat, nsamp = nsamp)
```

# model subset by sampling strategy
Determine how sampling strategy affects the relationships between the sim vars/nsamp

```{r fig.width=5, fig.height=4}
sampstrat_mod(df, sampstrat = sampstrat, full = TRUE)
```


```{r fig.width=5, fig.height=4}
nsamp_mod(df, nsamp = nsamp, full = TRUE)
```


# model subset by nsamp and sampling strategy
Determine how sampling strategy/nsamp affects the affects of the sim vars

```{r real code, fig.width=20, fig.height=4}
sampstrat_nsamp_mod_params(df, sampstrat = sampstrat, nsamp = nsamp, alpha = 0.10)
```


#FDRCOMBO
```{r fig.width=20, fig.height=8}
df$stat <- df$FDRCOMBO
summary_vplot(df)
summary_hplot(df)

moddf <- df[df$sampstrat != "full",]
```


# full model
```{r}
#Need to figure out if you need to subset this anova/remove n.s. variables at this stage or later
#mixed effect model
#na.action needs to be set to na.fail for dredge to work
fullmod <- lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + (1 | seed), moddf, na.action = "na.fail")

#anova
aovmod <- anova(fullmod)

#fixed effects (minus intercept)
efmod <- fixef(fullmod)[-1]

#print result
aovmod
efmod
```

```{r}
combos <- dredge(fullmod)
combos
```

```{r pairwise comparison of npts, fig.width=5, fig.height=5}

model <- lmer(stat ~ nsamp + K + m + phi + H + r + (1 | seed), moddf, na.action = "na.fail")

pw <- emmeans(model, list(pairwise ~ nsamp), adjust = "tukey")

resdf <- data.frame(pw$`pairwise differences of nsamp`)
resdf$p1 <- gsub("\\ -.*", "", resdf$X1)
resdf$p2 <- gsub(".*- ", "", resdf$X1)

resdf$estimate[resdf$p.value > 0.05] <- NA

resdf$p1 <- factor(resdf$p1, ordered=TRUE, levels = as.character(nsamp))
resdf$p2 <- factor(resdf$p2, ordered=TRUE, levels = as.character(nsamp))

ggplot(resdf, aes(p1,p2)) +
    geom_tile(aes(fill = estimate)) + 
    geom_text(aes(label = signif(estimate,  digits = 2), hjust = 0.5)) +
    scale_fill_gradient2(low = "#2066AC", mid="#F7F7F7", high = "#B2182B", midpoint = 0, limits=c(min(resdf$estimate),max(resdf$estimate))) +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          #legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 12),
          axis.text.y = element_text(color = "gray50", size = 12), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()

```



```{r pairwise comparison of sampstrat, fig.width=5, fig.height=5}
#TukeyHSD

model <- lmer(stat ~ nsamp + sampstrat + K + m + phi + H + r + (1 | seed), moddf, na.action = "na.fail")

pw <- emmeans(model, list(pairwise ~ sampstrat), adjust = "tukey")

resdf <- data.frame(pw$`pairwise differences of sampstrat`)
resdf$p1 <- gsub("\\ -.*", "", resdf$X1)
resdf$p2 <- gsub(".*- ", "", resdf$X1)


resdf$estimate[resdf$p.value > 0.05] <- NA

resdf$p1 <- factor(resdf$p1, ordered=TRUE, levels = as.character(c("envgeo","grid","rand","trans")))
resdf$p2 <- factor(resdf$p2, ordered=TRUE, levels = as.character(c("envgeo","grid","rand","trans")))

ggplot(resdf, aes(p1,p2)) +
    geom_tile(aes(fill = estimate)) + 
    geom_text(aes(label = signif(estimate,  digits = 2), hjust = 0.5)) +
    scale_fill_gradient2(low = "#2066AC", mid="#F7F7F7", high = "#B2182B", midpoint = 0, limits=c(min(resdf$estimate),max(resdf$estimate))) +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          #legend.position = "none",
          axis.title.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.ticks.y=element_blank(),
          axis.text.x = element_text(color = "grey50", size = 12),
          axis.text.y = element_text(color = "gray50", size = 12), 
          plot.margin=unit(rep(0.4,4),"cm")) +
    coord_fixed()

```

#which is best when?
```{r, fig.height=8, fig.width=5}
nsamp_sampstrat_mods(df, sampstrat = sampstrat, nsamp = nsamp)
```

```{r fig.width=10, fig.height=10}
sampstrat_nsamp_mods(df, sampstrat = sampstrat, nsamp = nsamp)
```

# model subset by sampling strategy
Determine how sampling strategy affects the relationships between the sim vars/nsamp

```{r fig.width=5, fig.height=4}
sampstrat_mod(df, sampstrat = sampstrat, full = TRUE)
```

```{r fig.width=5, fig.height=4}
nsamp_mod(df, nsamp = nsamp, full = TRUE)
```
# model subset by nsamp and sampling strategy
Determine how sampling strategy/nsamp affects the affects of the sim vars

```{r real code, fig.width=20, fig.height=4}
sampstrat_nsamp_mod_params(df, sampstrat = sampstrat, nsamp = nsamp, alpha = 0.10)
```

