---
title: "Optimizing sampling design for landscape genomics: Paramaterizing correlations between environmental layers"
author: "Anusha P. Bishop, Drew E. Terasaki Hart, Ian J. Wang"
output: html_document
---

To determine how to set the levels of correlation between our environmental layers we created distributions of correlations between temperature and precipitation globally. We downloaded global average temperature and precipitation data for 1970-2000 at a resolution of 2.5 minutes of a degree from WorldClim using the "geodata" package (Hijmans  et al., 2023). This data is monthly, so we took the mean across all months to get one raster layer representing temperature and one representing precipitation. To generate our distributions we randomly sampled one thousand 100 by 100 cell sections of each raster and calculated the correlation between temperature and precipitation within those sections. We took the absolute value of the correlations since for our simulations only the magnitude matters, not the direction. For our simulations we chose 0.30 for our low level and 0.60 for our high level which corresponded to approximately the first and second quartile (i.e., the median) of the distribution of correlations. We didn't select the third quartile value of 0.84 because when a pair of environmental variables has a correlation greater than 0.70 one of them is usually dropped in landscape genomic analyses. Therefore, we wanted our high level of correlation to still be below the threshold of being too correlated to be disentangled. The code used to create these distributions can be found in the same notebook used to generate this file [here](https://github.com/AnushaPB/LandGenSamp/blob/main/p1_gnxsims/MNLM/File_S2.Rmd).

```{r, include = FALSE}
library(terra)
library(tidyverse)
library(geodata)
library(here)
```

```{r, result = FALSE, message = FALSE, include = FALSE, cache = TRUE}
# Download temperature and precipitation data (tavg) at 2.5 arc-min resolution
tavg <- worldclim_global(var = "tavg", res = 2.5, path = here("p1_gnxsims", "MNLM"))
prec <- worldclim_global(var = "prec", res = 2.5, path = here("p1_gnxsims", "MNLM"))

tavg <- mean(tavg)
prec <- mean(prec)
env <- c(tavg, prec)
```


```{r fig.width = 10, fig.height = 5, include = FALSE, cache = TRUE}
# Create distribution of correlations
# x - raster stack of two layers
# ldim - subraster dimensions (creates a 100x100 subsampled raster)
# percent.na - maximum percentage of NA values allowed in the subsampled raster
# nsamp - how many subsampled rasters to generate
rcor <- function(x,
                 ldim = 100,
                 nsamp = 1000, 
                 percent.na = 0){
  
  #convert from pixels to coordinate units (assumes x and y units are the same)
  ldimconv <- ldim*res(x)[1]
  
  #Create buffered raster based on ldim (dimensions of subrasters)
  #so that when the subrasters are created in the next step they don't extend outside of the extent 
  #(only needed for xmax/ymax since the box is created by adding to an initial coordinate
  buffrast <- crop(x, ext(ext(x)[1],
                          ext(x)[2] - ldimconv,
                          ext(x)[3],
                          ext(x)[4] - ldimconv))
  
  #df for saving correlation
  cors <- c()
  
  #vector to save initial cell values so the same subraster isn't sampled twice
  initvec <- c()
  
  #get non na cell numbers
  notna <- which(complete.cases(values(buffrast)))
  
  #loop to create subrasters
  for(i in 1:nsamp){
    
    #Initial conditions for while loops
    pna <- 1
	  tn <- 0
	  init <- notna[1]
	  
	  #Loop through subrasters while the percent of NAs is above a threshold 
	  #AND while the number of attempts to get the raster to meet this condition is below 1000 (just so it doesn't run forever)
	  testn = 1000
  	while (pna > percent.na & tn <= testn){
  	  #Loop through initial points for the subraster corner that have not been sampled before
  	  while (init %in% initvec){
  	    init <- sample(notna, size = 1)
  	  }
  	  
  	  #add point to vector so it isn't sampled again
  	  initvec <- c(initvec, init)
  	  #get coordinates of the initial corner (lower left)
  	  coord1 <- xyFromCell(buffrast, init)
  	  #add ldim to corner to get the upper right corner of the subraster
  	  coord2 <- coord1 + ldimconv
  	  #create coords object for cropping
    	coords <- c(coord1[1], coord2[1], coord1[2], coord2[2])
    	#crop raster with coords to get subraster
  	  newx <- crop(x, ext(coords))
  	  #calculate the percentage of NAs
  	  pna <- sum(is.na(values(newx)))/(ncell(newx)*nlyr(x))
  	  #tracker for number of loops
  	  tn <- tn + 1
  	}
	  if (tn == testn){"failed to find a subraster below the NA threshold in testn attempts"}
	  #calculate and save pearson's correlation for each subraster
	  #calc correlation
	  cor <- layerCor(newx, "pearson", na.rm=T)$`pearson`[1,2]
	  cors <- append(cors, cor)
  }
	return(cors)
}
```

```{r, include = FALSE, cache = TRUE, include = FALSE}
set.seed(070085)
cors <- rcor(env)
```

```{r, fig.width = 9, fig.height = 3, echo = FALSE}
cors_df <-  
  data.frame(r = cors) %>%
  # get absolute value of r since we only care about magnitude
  mutate(r = abs(r))

ggplot(cors_df) +
  geom_boxplot(aes(x = r), fill = "cornflowerblue", alpha = 0.5, lwd = 1) +
  theme_classic() +
  geom_vline(aes(xintercept = 0.3), col = "aquamarine2", lwd = 1.5, lty = "dashed") +
  geom_vline(aes(xintercept = 0.6), col = "orange", lwd = 1.5, lty = "dashed") +
  ylim(-1,1) +
  theme(axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 18)) +
  xlab("|Correlation between Temperature and Precipitation|")
```

**Figure 1.** Boxplot of the absolute value of correlations between global temperature and precipitation. Green and orange dashed lines represent the low and high levels of correlation used in our simulation.

### References:

1. Fick, S.E. and R.J. Hijmans, 2017. WorldClim 2: new 1km spatial resolution climate surfaces for global land areas. International Journal of Climatology 37 (12): 4302-4315.

2. Hijmans RJ, Barbosa M, Ghosh A, Mandel A (2023). geodata: Download Geographic
  Data. R package version 0.5-8, <https://CRAN.R-project.org/package=geodata>.